Model_CFU

### 2016-12-16 Projects/HumanCdGF_cdiff_colonization.Rmd
rearrange Rmd and clean up language

@@

Removing the repeated cages of DA00581 (A and C) causes the model to perform poorly leaving one out.
So is the performance of the whole model appears to be dependent on two grouping of points (one around ~5-6 log and one around ~8 log ). So how well can random forest model the data with out the DA00581 source?

@@

When using all OTUs the model preforms poorly. The median Rsq of the trainset was `r median(om_rsq$train_rsq)`, this is likely do to the features. This is a summary of the importance scores for all predictors:  
`r summary(unlist(select(om_import_otus, -otu)))`  
`r round(100*sum(om_import_otus[,-1]<=0, na.rm = T)/length(om_import_otus[,-1]<=0),2)`% of OTUs have an importance of 0 or less. The median Rsq of the test set was `r median(om_rsq$test_rsq)` and the Rsq of all predicted values was `r om_rsq_overall`.

Looking at the % var explained values from each individual iteration, there appears to be a bimodal distribution in the performance.

Looking at the % var explained values from each individual iteration, there appears to be a bimodal distribution in the performance.

By cage:  
Min. 1st Qu.  Median    Mean 3rd Qu.    Max.   
`r summary(om_rsq_cage_df$rsq[om_rsq_cage_df$data_set=='test'])`

By donor:  
Min. 1st Qu.  Median    Mean 3rd Qu.    Max.

`r summary(om_rsq_pos_df$rsq[om_rsq_pos_df$data_set=='test'])`

```{r compare number of 581 donor cages, fig.height=5, fig.width=5}

rbind(data.frame(rsq = om_rsq_pos_df$rsq[om_rsq_pos_df$data_set=='test'], data = 'all'),
      rbind(data.frame(rsq = om_rsq_AD_df$rsq[om_rsq_AD_df$data_set=='test'], data = 'Two_only'),
            data.frame(rsq = om_rsq_C_only_df$rsq[om_rsq_C_only_df$data_set=='test'], data = 'One_only'))) %>% 
  ggplot(aes(rsq, color = data)) + 
  geom_density() + ggtitle('Rsq Distribution when multiple cages from same Donor present')
```

It seemed possible the assumption all donors have the same community may be false. However, when running the analysis with only one cage from the donor with multiple cages it performs as we observe for the lower peak in the bimodal distribution and when a second cage from the same donor is added we observe Rsq at the higher peak in the bimodal distribution. Thus there appears to be a cage effect improving the performance and the bimodal distribution maybe due to a variation in the ability of these communities to be able to predict each other, ie some be more similar to each other but still as a group more different than the other cages.

>>>>

When using all OTUs the model preforms poorly. The median Rsq of the trainset was `r median(om_rsq$train_rsq)`, this is likely do to the features. This is a summary of the importance scores for all predictors:  
Min.  1st Qu.   Median     Mean  3rd Qu.     Max.     NA's   
`r summary(unlist(select(om_import_otus, -otu)))`  
`r round(100*sum(om_import_otus[,-1]<=0, na.rm = T)/length(om_import_otus[,-1]<=0),2)`% of OTUs have an importance of 0 or less. The median Rsq of the test set was `r median(om_rsq$test_rsq)` and the Rsq of all predicted values was `r om_rsq_overall`.

The % var explained values from each individual iteration appears to have a bimodal distribution.

The distribution of Rsq values when modeling by cage appears to be improved in comparison of by human donor

By cage:  
Min., 1st Qu.,  Median,    Mean, 3rd Qu.,    Max.,   
`r summary(om_rsq_cage_df$rsq[om_rsq_cage_df$data_set=='test'])`

By donor:  
Min., 1st Qu.,  Median,    Mean, 3rd Qu.,    Max.,   
`r summary(om_rsq_pos_df$rsq[om_rsq_pos_df$data_set=='test'])`

```{r compare number of 581 donor cages, fig.height=5, fig.width=6}

rbind(data.frame(rsq = om_rsq_pos_df$rsq[om_rsq_pos_df$data_set=='test'], data = '3'),
      rbind(data.frame(rsq = om_rsq_AD_df$rsq[om_rsq_AD_df$data_set=='test'], data = '2'),
            data.frame(rsq = om_rsq_C_only_df$rsq[om_rsq_C_only_df$data_set=='test'], data = '1'))) %>% 
  ggplot(aes(rsq, color = data)) + 
  geom_density() + ggtitle('Rsq Distribution varying number of Donor 581 cages')
```

It seems the assumption all donors have the same community may be false because the bimodal distribution appears to be attributed to which mice from that group are used. However, when running the analysis with only one cage from the donor with multiple cages (581) it performs as we observe for the lower peak in the bimodal distribution and when a second cage from the same donor is added we observe Rsq values at the higher peak in the bimodal distribution. Thus there appears to be a cage effect improving the performance and the bimodal distribution maybe due to a variation in the ability of these communities to be able to predict each other, ie some may be more similar to each other but still as a group more different than the other cages.

## So how does this cage/donor affect the overall CFU prediction?

@@

Removing the repeated cages of DA00581 (A and C) causes the model to perform poorly leaving one out.  
The performance of the whole model appears to be dependent on two grouping of points (one around ~5-6 log (all Donor 581) and one around ~8 log ).  
__How well can random forest model the data with out the DA00581 source?__

@@
